Sparse spectral and $p$-finite element methods on disk-slices and trapeziums I: zero boundary conditions

### 2019-03-27 papers/halfdisk/halfdisk.tex
Further edits to paper. Included updated figures

We consider the solution of partial differential equations on the domain
\begin{align}
\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align}
For simplicity of discussion we focus on the half-disk, where $\rho(x) = \sqrt{1-x^2}$,  $(\alpha,\beta) = (0, 1)$, and  $(\gamma, \delta)  = (-1,1)$, with discussion of extension to other geometries in the appendix.

We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the triangle \cite{olver2018recurrence,olver2019triangle}.  On the half-disk the family of weights we consider are of the form
$$
W^{(a,b)}(x,y) = x^a (1-x-y)^b
$$
with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$. We define these to be orthogonalised lexicographically, that is,
$$
P_{n,k}^{(a,b)}(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
$$
where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.

@@

The motivation for this work is to solve surface PDEs on the sphere.
In subsequent work we will extend  directly by projecting up to 3D OPs on the hemisphere, where sparse operators, function evaluation and expansion etc. can be found similarly.

@@

In this section we outline some basic properties of $P_{n,k}^{(a,b)}(x,y)$. The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.

@@

\begin{proposition}
Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\),
Then, \(\forall n,k = 0,1,2,\dots\) let \(\{p_{n,k}\}\) be polynomials orthogonal with respect to the weight \(\rho(x)^{2k+1} w_1(x)\), and \(\{q_{n}\}\) be polynomials orthogonal with respect to the weight \(w_2(x)\). Then
$$
\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
$$
are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2(\frac{y}{\rho(x)}) \) on $\Omega$. 
\end{proposition}

For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:

\begin{align}
\begin{cases}
(\alpha,\beta) &:= (0.1) \\
(\gamma,\delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= x^a \: (1-x^2)^b \\
w_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b,
\end{cases}
\end{align}

The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\). The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials. Thus we arrive at  a two parameter family of  2D orthogonal polynomials \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align}
 \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on \(\Omega\) with respect to the weight $W^{(a,b)}(x,y)$.

@@

Express the three-term recurrences associated to $H_n^{(a,b)}$ and $P_n^{(b,b)}$ as
\begin{align}
x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x) \\
y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
\end{align}
where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $P_{n,k}^{(a,b)}(x,y)$:

\begin{lemma}
$\Pnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
\begin{align}
x \Pnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: P_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: P_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: P_{n+1, k}^{(a,b)}(x, y), \\
y \Pnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: P_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: P_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,3} \: P_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: P_{n, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,5} \: P_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: P_{n+1, k+1}^{(a,b)}(x, y),
\end{align}
for \((x,y) \in \Omega\), where
\begin{align}
\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
\end{align}
where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{H^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight.

@@

\begin{proof}
Note that \(\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
\begin{align}
\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &=  \iint_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
&= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
&= \begin{cases}
    	\delta_k^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
	\delta_{k-1}^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
	0 & \text{otherwise}
      \end{cases}.
\end{align}

@@

Further, we note that there exist "parameter transformation" Banded-Block-Banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another. We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T^{(a,b)\to(a-1,b)}$, $T^{(a,b)\to(a,b-1)}$ and $T^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
\begin{align}
\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b)} \Big)^T \: \bigP^{(a+1,b)}(x,y) \\
\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigP^{(a+1,b+1)}(x,y) \\
\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b)} \Big)^T \: \bigW^{(a-1,b)}(x,y) \\
\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a,b+1)} \Big)^T \: \bigW^{(a,b-1)}(x,y) \\
\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigW^{(a-1,b-1)}(x,y) \\
\end{align}

@@

Denote the  Gauss quadrature nodes and weight on \([0,1]\) with weight \(s^a \: (1-s^2)^{b+\half}\) as $(s_k,w_k^{(s)})$ , and
 on \([-1,1]\) with weight \((1-t^2)^b\) as $(t_k,w_k^{(t)})$. Define
\begin{align}
x_{i+(k-1)N} &:= s_k, \quad i,k = 1,\dots,N, \\
y_{l+(i-1)N} &:= (1-s_l^2)^\half \: t_l, \quad i,l = 1,\dots,N, \\
w_{l+(k-1)N} &:= w_k^{(s)} w_l^{(t)}, \quad k,l = 1,\dots,N.
\end{align}
If $f(x,y)$ is an at most degree $N$ polynomial in $x$ and at most degree $2N-1$ in $y$ then
$$
\iint_\Omega f(x,y) \: \Wab(x,y) \: \D A = \half \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
$$

\end{theorem}
\begin{proof}

We will use the substitution that
\begin{align}
x &= s, \quad y = (1-s^2)^\half t.
\end{align}
First, note that
\begin{align}
W^{(a,b)}(x,y) &= x^a \: (1-x^2-y^2)^b, \quad \text{for } (x,y) \in \Omega, \\
		      &= s^a \: (1-s^2)^{b} \: (1-t^2)^b =: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
\end{align}

@@

\bstodo{This section contains repetition / expansion on the quadrature rule section above - not sure if we should remove it or move it?} We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficient vectors \(\mathbf{f}_n\) as follows.

Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$ (i.e. $f(\cdot, y)$ with $y$ fixed is a degree $d_1$ polynomial, and $f(x, \cdot)$ with $x$ fixed is an even degree $d_2$ polynomial). Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e. that $d_2$ is an even integer). Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree $d_1+\frac{d_2}{2}$ in $s$ and degree $d_2$ in $t$. Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
\begin{align}
& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
\end{align}
where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.

We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
\begin{align}
\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
\end{align}

Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
\begin{align}
\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
\end{align}
and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$. Note that for a general polynomial where the overall degree is $N$, we still would hence need $N^2$ points and weights to make the rule exact. Thus,
\begin{align}
\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big],
\end{align}
for suitably large $N$.

>>>>

We consider the solution of partial differential equations on the domain
\begin{align}
\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}.
\end{align}
For simplicity of discussion we focus on the half-disk, where $\rho(x) = \sqrt{1-x^2}$,  $(\alpha,\beta) = (0, 1)$, and  $(\gamma, \delta)  = (-1,1)$, with discussion of extension to other geometries in the appendix.

We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the triangle \cite{olver2018recurrence,olver2019triangle}.  On the half-disk the family of weights we consider are of the form
$$
W^{(a,b)}(x,y) = x^a (1-x-y)^b
$$
with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
$$
P_{n,k}^{(a,b)}(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
$$
where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.

@@

@@

In this section we outline the construction and some basic properties of $P_{n,k}^{(a,b)}(x,y)$. The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.

@@

\begin{proposition}
Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and let \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
\begin{enumerate}
\item  \(\rho\) is a degree 1 polynomial, or 
\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
\end{enumerate}
$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$. Then the 2D polynomials defined on $\Omega$
$$
\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
$$
are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2(\frac{y}{\rho(x)}) \) on $\Omega$. 
\end{proposition}
\begin{proof}
See \cite[p55--56]{dunkl2014orthogonal}.
\end{proof}

For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
\begin{align}
\begin{cases}
(\alpha,\beta) &:= (0.1) \\
(\gamma,\delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= w_H^{(a,b)}(x) := x^a \: (1-x^2)^b \\
w_2(x) &:= w_P^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
\end{cases}
\end{align}

The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\). The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials. Thus we arrive at  a two parameter family of  2D orthogonal polynomials \(\{\Pnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align}
 \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on \(\Omega\) with respect to the weight
\begin{align}
W^{(a,b)}(x,y) = w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
\end{align}

@@

We can express the three-term recurrences associated to $H_n^{(a,b)}$ and $P_n^{(b,b)}$ as
\begin{align}
x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x) \\
y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
\end{align}
where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $P_{n,k}^{(a,b)}(x,y)$:

\begin{lemma}
$\Pnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
\begin{align}
x \Pnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: P_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: P_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: P_{n+1, k}^{(a,b)}(x, y), \\
y \Pnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: P_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: P_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,3} \: P_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: P_{n, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,5} \: P_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: P_{n+1, k+1}^{(a,b)}(x, y),
\end{align}
for \((x,y) \in \Omega\), where
\begin{align}
\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}}. 
\end{align}
where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{w_H^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight of $w_H^{(a, b)}$.

@@

\begin{proof}
Note that \(\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
\begin{align}
\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &=  \iint_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
&= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
&= \begin{cases}
    	\delta_k^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
	\delta_{k-1}^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
	0 & \text{otherwise}
      \end{cases}.
\end{align}

@@

Further, we note that there exist "parameter transformation" Banded-Block-Banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another. We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
\begin{align}
\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b)} \Big)^T \: \bigP^{(a+1,b)}(x,y) \\
\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigP^{(a+1,b+1)}(x,y) \\
\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b)} \Big)^T \: \bigW^{(a-1,b)}(x,y) \\
\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a,b+1)} \Big)^T \: \bigW^{(a,b-1)}(x,y) \\
\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigW^{(a-1,b-1)}(x,y) \\
\end{align}

@@

Denote the  Gauss quadrature nodes and weight on \([0,1]\) with weight \(s^a \: (1-s^2)^{b+\half}\) as $(s_k,w_k^{(s)})$ , and
 on \([-1,1]\) with weight \((1-t^2)^b\) as $(t_k,w_k^{(t)})$. Define
\begin{align}
x_{i+(k-1)N} &:= s_k, \quad i,k = 1,\dots,N, \\
y_{l+(i-1)N} &:= (1-s_l^2)^\half \: t_l, \quad i,l = 1,\dots,N, \\
w_{l+(k-1)N} &:= w_k^{(s)} w_l^{(t)}, \quad k,l = 1,\dots,N.
\end{align}
Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
$$
\iint_\Omega f(x,y) \: \Wab(x,y) \: \D A \approx \half \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
$$
where if $f(\cdot, y)$ for fixed $y$ is an at most degree $N$ polynomial and $f(x, \cdot)$ for fixed $x$  is an at most degree $2N-1$ polynomial, the quadrature rule is exact.
\end{theorem}

\begin{proof}
We will use the substitution that
\begin{align}
x &= s, \quad y = (1-s^2)^\half t.
\end{align}
First, note that
\begin{align}
W^{(a,b)}(x,y) &= x^a \: (1-x^2-y^2)^b, \quad \text{for } (x,y) \in \Omega, \\
		      &= s^a \: (1-s^2)^{b} \: (1-t^2)^b =: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
\end{align}
\end{proof}

@@

Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$:
\begin{align}
\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big].
\end{align}
